### 2021-07-29 paper-draft/paper.tex
Added backup of paper

\documentclass{article}
\usepackage[backend=biber, style=authoryear-icomp]{biblatex}
\usepackage{blindtext}
\usepackage{multicol}
\usepackage[document]{ragged2e}
\usepackage[pass, letterpaper]{geometry}
%\usepackage[margin=1in]{geometry}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{subcaption}
\addbibresource{ref.bib}

% Define macros
\newcommand\note[1]{\textbf{\textcolor{red}{#1}}}

% Captions
\captionsetup{justification = raggedright, singlelinecheck = false, labelfont=bf}

\title{Using a Deep Autoencoder Model to Characterize Air Pollution}
\author{Nicolas Shannon$^{1}$, Bishnu Timalsena$^{2}$, Raul Chavez Jr.$^{3}$, \\Vandana Nunna Lakshmi$^{2}$}
\date{}
% To cite, use \parencite{} or \textcite{}

\begin{document}

% Acknowledgements
\begin{centering}
\maketitle
\vspace{-0.50cm}
\scriptsize
{$^{1}$Department of Computer Science, Loyola University Chicago, Chicago, IL, US}\\
{$^{2}$Department of Computer Science, University North Texas, Denton, TX, US}\\
{$^{3}$Department of Electrical and Computer Engineering, University of Texas 
    - Rio Grande Valley, Brownsville, TX, US \note{$\leftarrow$ fix formatting}}
\end{centering}

\begin{abstract}
Air pollutants are both a cause and a symptom of climate change. Global warming has the potential to raise
the levels of various air pollutants. The rise in surface air pollutants can lead to heat trapped close to 
the Earth's surface by virtue of the greenhouse effect, which further perpetuates the cycle. In order to
more effectively combat climate change, it is important to understand and characterize the source of 
emissions. We developed a technique to characterize cities by the levels of various air toxins using an
undercomplete autoencoder model. By representing the pollution levels of eighteen thousand cities across
the United States, we explore some of the more prominent features that characterize urban pollution.
We compare two techniques for reducing the input dimensionality - Principle Component Analysis (PCA) and
Undercomplete Autoencoders.
\end{abstract}

\vspace{0.25cm}
\section{Introduction}
Over the past several decades, air pollution has been a major concern for human health
and the environment. The EPA sets national air quality standards for  six major air
toxins: ground level Ozone (O$_3$), Carbon Monoxide (CO), Sulfur Dioxide (SO$_2$),
Lead (Pb), Nitrogen Dioxide (NO$_2$), and particulate matter (PM). \parencite{NAAQS09}.
The sources of these emissions vary, but the two largest contributors are industrial 
processes and motorized vehicles.

\newpage

Short and long term exposure to air toxins pose significant risk to human health. 
A study by \textcite{Wang19} found a correlation  between increased ambient CO
levels and outpatient visits for respiratory, cardiovascular, genitourinary, and gastrointenstinal
diseases. Several studies performed by the EPA also suggest that people with pre-existing health
conditions are at higher risk when exposed to CO over long periods.\par

\subsection{Air Toxins}
We wanted to explore how CO and other gases can be characterized with respect to location
in an effort to provide researchers and policy makers with the context and tools necessary to
make informed decisions. There are a litany of technologies and tools for pollution prevention
and control, but most of these options require contextual information and large amounts of data.
\par
By creating an embedding of time series pollution data for numerous cities throughout the U.S.,
we can explore the more interesting features of the model as well as apply it to tangential 
areas such as the pollution generated by industrial farming and animal husbandry activities.

\begin{figure}[h]
\includegraphics[width=\textwidth]{/home/nicks/Documents/paper-draft/images/co-concentration-table.png}
    \caption{Carbon Monoxide Exposure Table, Parts Per Million (ppm)}
    \label{fig:co-exposure}
\end{figure}

\subsection{A Deep Learning Approach}
In this paper we compare the autoencoder model with principle component analysis (PCA), a much
older technique for dimensional reduction. Using a deep autoencoder has several advantages.
First, deep autoencoders are better at compressing highly complex data into few dimensions.
Autoencoders with a nonlinear encoder and decoder function can learn more powerful nonlinear
generalizations than PCA. \parencite{Goodfellow16}.
\par \note{Autoencoder grapic from poster here}

\section{Related Work}
The idea of generating embeddings for a location is a highly potential expanse for research.
The two major forces behind its competency are, the fact that large volumes of data is being
effectively compressed and that these embeddings, when plugged into any model could effectively
represent any given location.  A lot of research is being done on location embeddings in recent
times. For instance, location embeddings are being generated based on human-mobility data 
(Crivellari, A.; 2019) to understand the behavioral proximity between different locations 
irrespective of their spatial distances. The sequence of locations that are frequently
travelled were fed to a Skip-gram Word2vec model to generate the location embeddings. Another 
approach that was implemented to leverage location embeddings is, GPS2Vec ( Yifang Yin, 2019).
This application consists of a neural network that would generate embeddings for a location, 
based on the semantic contexts that are sourced from the user content published on multiple 
digital platforms. With the integration of these embeddings, a successful geo-tagged image
classification task is demonstrated. In another effort (Carl Yang, 2019), the location 
embeddings derived from user content such as addresses, phone numbers, images, place names and 
all other details from multiple individual online posts, are compressed to form low dimensional
embeddings to solve the place duplication issue. Deduplication of locations, is a paramount 
task for any digital platform since they provide a place graph to its users, which is the 
result of merging location data from multiple sources. This merge leads to duplicity since 
different sources could hold different names for the same place. By applying K-NN search, 
pair-wise duplication prediction on these generated location embeddings, the application can
identify the duplicates and eliminate them to result in an efficient place graph.

\section{Methodology}

We used two different methods to reduce the dimensionality of our data - Principle Component
Analysis and Undercomplete Autoencoders. A simple linear regression was performed on 
the encoded data that was generated by both the PCA and AE models. Additionally, we performed
a linear regression on the unencoded values for comparison. Each model's goodness of fit
was evaluated using the percentage of explained variance for the three linear regressions.

% pipeline
\begin{figure}[h]
\includegraphics[width=\textwidth, height=6cm]{/home/nicks/Documents/paper-draft/images/pipeline.png}
\caption{Data Pipeline}
\label{fig:pipeline}
\end{figure}

\subsection{Data Collection and Preprocessing}
The dataset was created using OpenWeather's Air Pollution API. The dataset includes data for 
eight polluting gases and particulates: CO, NO, NO$_2$, O$_3$, SO$_2$, NH$_3$, PM$_{2.5}$, 
and PM$_{10}$. The dataset also includes the hourly timestamp, the latitude and longitude (part of the input),
and the air quality index, the latter of which we did not end up using.\par
We initially looked at data for the eight pollutants over a six-and-a-half month time 
frame: January 27th, 2020 through July 6th, 2021. However, we found that we achieved
better results with just four months of of data for our input features. Our final 
dataset included just over eighteen thousand cities in the United States and other U.S. 
territories. Additionally, we found it easier to focus on a single gas when developing 
the models, so we chose to concentrate on CO during the engineering process. Given
enough time and computational power, the techniques performed on the CO model would be 
easily extensible and transferable to the other seven pollutants.

% dataset characteristics table
\begin{table}[h!]
    \caption{Dataset Characteristics}
    \label{tab:table1}
    \vspace{0.1cm}
    \begin{tabular}{p{4cm}p{7cm}}
        \hline
        \multicolumn{2}{c}{OpenWeather Pollution API} \\
        \hline
        Time period  & 6.5 months    \\
        Time frequency & Daily  \\
        $\#$ of cities & 18,526 \\
        $\#$ of features & 120 (originally 190) \\
        $\#$ of component gases & 8 \\
        Dependent variable & Single day of time series (3/27/21) \\
        \hline
    \end{tabular}
\end{table}

OpenWeather's Pollution API ordinarily returns hourly values, but we opted to use the daily
average pollution levels as our feature set. We resolved to feature engineer the daily 
averages and to reduce the time scale to four months in order to keep the data 
proportional to the number of features and thus avoid the curse of dimensionality. 
\parencite{Trunk79}. \note{Idk about this} \par
We separated the dependent from the input before unit normalizing the input. 
The normalized data was then encoded by each model.

\subsection{Autoencoder and PCA Models}
The autoencoder model used consists of an encoded and decoded layer. We performed
a grid search for the CO model to determine the optimal hyperparameters. The following 
hyperparameters were tested:
\begin{itemize}
    \item \textit{Learning Rate:} 0.0001, 0.001, 0.01, 0.1
    \item \textit{Batch Size:} 32, 64, 128, 256
    \item \textit{Epochs:} 10, 50, 75, 100
\end{itemize}
We used the parameters that appeared most frequently over the one hundred twenty dimensions and 
applied them over the entire range so as not to overfit.\note{$\leftarrow$ not sure if this is correct}

\begin{table}[h!]
    \caption{Dataset Characteristics}
    \label{tab:table1}
    \vspace{0.1cm}
    \begin{tabular}{p{4cm}p{7cm}}
        \hline
        \multicolumn{2}{c}{Optimal Hyperparameters for Carbon Monoxide (CO)} \\
        \hline
        Activation function & Leaky ReLU \\
        Loss function & Mean Squared Error  \\
        Optimizer & Adam \\
        Learning rate & 0.001 \\
        Batch size & 64 \\
        Epochs & 150 \\ 
        \hline
    \end{tabular}
\end{table}

The PCA model was trained with k-fold cross-validation using five folds. The dataset was
split into a training, validation, and hold-out test set.

\section{Results}
First, we wanted to explore the correlation between days in our dataset. We calculated
the correlation coefficients and created a correlation matrix of the days that had
correlation coefficients of the highest magnitude.\par

% pipeline
\begin{figure}[h]
\includegraphics[width=\textwidth]{/home/nicks/Documents/paper-draft/images/corr_matrix_25.png}
\caption{Correlation Matrix}
\label{fig:pipeline}
\end{figure}

We compared the percent explained variance across one hundred twenty dimensions for
the encoded values generated by the PCA and autoencoder models as well as for the unencoded
values by using a simple linear regression. While PCA was overall less noisy, we found that
the autoencoder model achieved a much higher explainable variance for the first few dimensions.
\\ \note{Note that results graphs are placeholders}
% ae vs pca linegraph
\begin{figure}[h!]

\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=1.1\linewidth, height=4cm]{/home/nicks/Documents/paper-draft/images/ae_all_dims.png} 
    \caption{Autoencoder Dimensionality Reduction}
    \label{fig:ae_dim_reduction}
\end{subfigure}% <- % sign must be here!
\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=1.1\linewidth, height=4cm]{/home/nicks/Documents/paper-draft/images/ae_all_dims.png}
    \caption{PCA Dimensionality Reduction}
    \label{fig:pca_dim_reduction}
\end{subfigure}

\caption{Graph of autoencoder and PCA reduced representations of air pollutants}
\label{fig:ae_vs_pca}
\end{figure}

\par We clustered the cities based on the first two latent dimensions in order to see how
the model characterizes each city

% co scatter
\begin{figure}[h!]

\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=1.1\linewidth, height=5cm]{/home/nicks/Documents/paper-draft/images/co_scatter_ae_wlist.png} 
    \caption{Outlier Cities}
    \label{fig:outliers}
\end{subfigure}% <- % sign must be here!
\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=1.1\linewidth, height=5cm]{/home/nicks/Documents/paper-draft/images/co_scatter_ae_large_cities.png}
    \caption{Densely Populated Cities}
    \label{fig:dense_cities}
\end{subfigure}

\caption{Scatter plots of outlier and highly-populated cities over first two dimensions}
\label{fig:outliers_vs_dense_cities}
\end{figure}

By selecting from a list the top two hundred highly-populated cities, we found that nearly
all of them were located near the center of the clustering. Conversely, we found that many 
of the outliers were smaller cities in rural locations.

\section{Discussion}

\section{Conclusion}

\printbibliography

\end{document}